### 2024-03-01 HelmetEDADamping.Rmd
Add files via upload

---
title: "EDA-Helmet"
author: "Katie Greene"
date: "2024-02-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
foam_data <- read.csv(file.choose())

```

```{r}

ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=freq), colour = 'red') 
ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=amp), colour = 'orange')
ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=initialStrain), colour = 'green')
ggplot(data = foam_data, aes(y = damping)) + geom_point(aes(x=initialStress), colour = 'blue')
ggplot(data = foam_data, aes(x=cycle, y = damping , fill = factor(ID))) + geom_point()
```

```{r}
ggplot(data = foam_data, aes(x = porosity, y = damping, fill = factor(chemIndex))) + geom_boxplot()

```

```{r}

ggplot(foam_data, aes(x = factor(chemIndex), y = damping)) + geom_boxplot()
ggplot(foam_data, aes(x = factor(porosity), y = damping)) + geom_boxplot()
```

```{r}
ggplot(foam_data, aes(x = factor(cycle), y = damping, fill = factor(cycle))) + geom_boxplot()
```

```{r}
hist(foam_data$damping)
```

```{r}
ggplot(foam_data, aes(x = factor(ID), y = damping)) + geom_point()
tally(foam_data$damping ~ foam_data$ID)
```