### 2024-03-01 AvaEDA.Rmd
Add files via upload

---
title: "Ava's EDA"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*.

```{r}
library(ggplot2)
library(patchwork)
library(tidyverse)
```

```{r}
setwd("C:/Users/Ava/OneDrive - Colorado School of Mines/Desktop/Mines/Year4/Semester2/Stat Practicum")
foam_data <- read.csv("foamData.csv")
```

```{r}
ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = porosity)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = initialStrain)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = initialStress)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = chemIndex)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = freq)) +
  xlim(0,10000)

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = amp)) +
  xlim(0,10000)
```
```{r}
ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point() +
  xlim(0,10000)
```

```{r}
ggplot(foam_data, aes(x = porosity, y = stiffness)) +
  geom_boxplot()

ggplot(foam_data, aes(x = factor(chemIndex), y = stiffness)) +
  geom_boxplot()

ggplot(data= foam_data, aes(x = porosity, y = stiffness, fill = factor(chemIndex))) +
  geom_boxplot() 
```

```{r}
ggplot(foam_data, aes(x = cycle, y = damping)) +
  geom_point(aes(color = porosity))

ggplot(foam_data, aes(x = cycle, y = stiffness)) +
  geom_point(aes(color = initialStrain))
```