### 2024-04-09 mixedmodel.Rmd
Add files via upload

---
title: "mixed model"
output: html_document
date: "2024-04-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
setwd("C:/Users/Ava/OneDrive - Colorado School of Mines/Desktop/Mines/Year4/Semester2/Stat Practicum")
foam_data <- read.csv("foamData.csv")
```

```{r}
foam_data2 <- na.omit(foam_data)

foam_data3 <- filter(foam_data2$cycle<= 10000)
```

```{r}
head(foam_data2)
```
## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
library(lme4)
```

```{r}
model1 <- lmer(stiffnessFrac ~ freq + amp + chemIndex + porosity + porosity:chemIndex + log(cycle) + log(cycle):freq + log(cycle):amp + log(cycle):chemIndex + log(cycle):porosity + log(cycle):porosity:chemIndex + (1 + cycle|ID), data = foam_data2)
```