### 2024-04-18 finalModels.Rmd
Add files via upload

---
title: "Final Models"
author: "Yellow Team"
date: "2024-04-18"
output: pdf_document
---

```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE)
```

```{r}
library(lme4)
library(lmerTest)
library(leaps)
library(tidyverse)
```

## Data wrangling

```{r}
foam_data <- read.csv("HelmetFinal.csv")
```

```{r}
foam_data_scaled <- foam_data |>
  filter(cycle <= 10000 & cycle > 0)
```

## Damping Model

```{r}
regfit.full <- regsubsets(log(dampingFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + chemIndex:porosity + (chemIndex + porosity + freq + amp + initialStrain + initialStress + chemIndex:porosity)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

plot(regfit.full, scale = "bic")

```

**Theoretical Model**

$log(dampingFrac) = \alpha_0 + \alpha_1*I(porosity = low) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*freq + \beta_0*log(cycle) + \beta_1*(I(porosity = low)*log(cycle)) + \beta_3*(I(chemIndex = 2)*log(cycle)) + \beta_4*(I(chemIndex = 3)*log(cycle)) + \beta_5*(freq*log(cycle)) + u_i + v_i*log(cycle) + \epsilon_{ij}$

```{r}
dampModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

plot(dampModel)
qqnorm(resid(dampModel))
qqline(resid(dampModel))
```

```{r}
summary(dampModel)
```

```{r}
re <- ranef(dampModel)$ID

plot(re)
```

## Stiffness Model

```{r}
regfit.full <- regsubsets(log(stiffnessFrac) ~ chemIndex + porosity + freq + amp + initialStrain + initialStress + log(cycle) + (chemIndex + porosity + freq + amp + initialStrain + initialStress)*log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
#regfit.full <- regsubsets(stiffnessFrac ~ ., data = foam_data_scaled)
plot(regfit.full, scale = 'bic')
```

**Theoretical Model**

$log(stiffnessFrac) = \alpha_0 + \alpha_1*I(porosity = low)) + \alpha_2*I(chemIndex = 2) + \alpha_3*I(chemIndex = 3) + \alpha_4*I(chemIndex = 2)*I(porosity = low) + \alpha_5*I(chemIndex = 3)*I(porosity = low) + \beta_0*log(cycle) +  \beta_1*log(cycle)*I(porosity = low) + \beta_2*log(cycle)*I(chemIndex = 2) + \beta_3*log(cycle)*I(chemIndex = 2)*I(porosity = low) + \beta_4*log(cycle)*I(chemIndex = 3)*I(porosity = low) + u_i + v_i*log(cycle) + \epsilon_{ij}$

```{r}
FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) + factor(chemIndex):porosity +  log(cycle) +  factor(chemIndex):log(cycle) + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)
plot(FactorModel)
qqnorm(resid(FactorModel))
qqline(resid(FactorModel))
```

```{r}
summary(FactorModel)
```

```{r}
re <- ranef(FactorModel)$ID

plot(re)
```