### 2024-05-06 CS2 Final Draft.qmd
Edit Appendix

---
title: "Helmet Foam Case Study Report"
author: 'Katie Greene, Ava Harman, Chris Seery, Ian Stonecypher'
date: 'May, 2024'
format: pdf
---

```{r}
#| echo: false
#| include: false
#| label: front-matter
#| warning: false
# clean up & set default chunk options
rm(list = ls())
knitr::opts_chunk$set(echo = FALSE)

# packages
library(tidyverse) # for example
library(ggplot2) # for example
library(car)       # for example
library(tinytex)   # for example
library(patchwork)
library(lme4)
library(lmerTest)
library(leaps)
foam_data <- read.csv("foamData.csv")

foam_data_scaled <- foam_data |>
  filter(cycle <= 10000 & cycle > 0) |>
  mutate(dampingFrac = replace(dampingFrac, dampingFrac < 0, NA))

FactorModel <- lmer(log(stiffnessFrac) ~  factor(chemIndex) + freq + amp +  log(cycle) +  factor(chemIndex):log(cycle) + freq:log(cycle) + amp:log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

dampingModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) + (1 + log(cycle)|ID), data = foam_data_scaled)

# Convert intervalCode to a factor with specified levels
foam_data_scaled$intervalCode <- factor(foam_data_scaled$intervalCode, levels = c("A", "B", "C", "D","E", "F", "G", "H","I", "J", "K", "L","M", "N"))

```

## 1. Project Description

This project comes to us from Moira Foster, from the Lawrence Livermore National Laboratory. Foster has been running experiments on the Colorado School of Mines campus to study the fatigue responses of stiffness and damping in different varieties of foam. In her designed experiment, pieces of foam were put into a universal load frame, which repeatedly compressed and decompressed each piece for thousands of cycles, and each piece had its stiffness and damping measured periodically. These measurements, as well as the parameters of each experimental run, were given to us to aid in the experiment’s analysis stage.

Foster’s objective is to understand what affects the changes in stiffness and damping in a piece of foam under repeated light compressions. The goal of our analysis is to discover patterns and build models to provide this understanding. With this understanding, the U.S. Military could improve the foam padding in marine special combat helmets, ideally so they can survive at least 10 years of regular use.

## 1.1 Research Questions

**Question 1:** What parameters are impacting the stiffness and damping of foam used in military helmets?

**Question 2:** How are these parameters impacting the fatigue response in stiffness and damping over time?

## 1.2 Variables

\begin{center}
\textbf{Explanatory Variables:}
\end{center}

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Variable Name} & \textbf{Description} \\
\hline
ID & The unique identifier of the different foam pieces \\
chemIndex & The chemical makeup type of the foam. 1 = 79, 2 = 100, 3 = 121 \\
porosity & The porosity level of the foam. Low = 71\%, high = 81\% \\
freq & The number of compressions per second (hz) \\
amp & The amplitude of the compressions \\
initialStrain & The initial measure of the strain of foam. (Originally labeled ``strain'') \\
initialStress & The initial measure of the stress of the foam. (Originally labeled ``stresskpa'') \\
cycle & The cycle number the measurement was taken at \\
\hline
\end{tabular}
\end{center}

\begin{center}
\text{Table 1: Explanatory Variables}
\end{center}

\begin{center}
\textbf{Response Variables}
\end{center}

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Variable Name} & \textbf{Description} \\
\hline
stiffness & Measurement of the foam’s stiffness. (Originally labeled ``E1\_MPA'') \\
damping & Measurement of the foam’s damping. (Originally labeled ``delta\_rad'') \\
stiffnessFrac & Stiffness / first stiffness measurement \\
dampingFrac & tan(damping) / tan(first damping measurement) \\
\hline
\end{tabular}
\end{center}

\begin{center}
\text{Table 2: Response Variables}
\end{center}

The formulas for stiffnessFrac and dampingFrac were specified by Foster. She believed these would be more useful alternatives to studying raw stiffness and damping.

#  2. Exploratory Data Analysis (EDA)

We reformatted the data from wide to long to ensure each observation represented one measurement, not one foam sample. Measurements taken beyond 10,000 cycles were removed to maintain consistency across observations, and because Moira identified this as the most relevant time frame.

There were a few concerning observations with negative values of stiffness, likely due to a fault in the measuring machine. At Foster’s instruction, we converted these negative stiffness values to positive. Then we made the transformed response variable stiffnessFrac, which reports each stiffness measurement divided by that foam sample's initial stiffness measurement.

There were many observations with negative values of damping, but we were instructed to leave these as they were. Then we made the transformed response variable dampingFrac, which reports the tangent of each damping measurement divided by the tangent of that foam sample's initial damping measurement. DampingFrac had five data points that were drastic outliers, but these were changed to NA per Dr. Philtron's recommendation.

Printed below are how many of each type of foam sample was tested:

| Porosity | **ChemIndex 79** | **ChemIndex 100** | **ChemIndex 121** |
|---------:|:-----------------|:------------------|:------------------|
| **81**   | 9                | 7                 | 6                 |
| **71**   | 11               | 5                 | 8                 |

\begin{center}
\text{Table 3: Foam Sample Type Counts}
\end{center}

These counts show that a relatively small number of each type of foam were tested. The amounts are unequal as well, ranging from 5 to 11.

```{r}
#| echo: false
#| label: fig-1
#| warning: false
#| fig-cap: "Histogram of Frequencies Tested"
#| fig-align: "center"
#| fig-width: 4
#| fig-height: 4

foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", "initialStress")]
foam_Samples <- unique(foam_Samples)

ggplot(foam_Samples, aes(x=freq)) +
  geom_histogram(bins=10)
```

The histogram in @fig-1 shows the distribution of compression frequencies among the foam samples revealing that certain compression frequencies were tested more frequently than others. These non-uniformities in sample distribution might impact the reliability of conclusions drawn from the analysis.

```{r}
#| echo: false
#| label: fig-2
#| warning: false
#| fig-cap: "StiffnessFrac and DampingFrac vs Cycle"
#| fig-align: "center"
#| fig-width: 10
#| fig-height: 4

p1 <- ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle, col=factor(ID))) +
  geom_point() + 
  theme(legend.position="none")

p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle, col=factor(ID))) + 
  geom_point() +
  theme(legend.position="none")

combined <- p1 + p2

combined
```
In order to see how stiffnessFrac and dampingFrac change over time for all foams, we plotted a scatter plot for each vs cycle. The dot colors in @fig-2 represent which foam sample they were measured from. As cycle number increases, the stiffnessFrac and dampingFrac values not only change, but spread further apart. StiffnessFrac and dampingFrac both appear to have a relationship with cycle, but that relationship looks different for each foam sample.

```{r}
#foam_data_interval_N <- foam_data_scaled[foam_data_scaled$intervalCode == "N", ]

#ggplot(foam_data_interval_N, aes(y = stiffnessFrac, x = cycle)) + geom_point()

#foam_data_scaled$intervalCode == "N"

```

```{r}
#| echo: false
#| label: fig-3
#| warning: false
#| fig-cap: "Log(stiffnessFrac) vs Log(cycle) and Log(dampingFrac) vs Log(cycle), Colored by Freq"
#| fig-align: "center"
#| fig-width: 10
#| fig-height: 5

sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = freq))
#sp2 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
#sp3 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) + geom_point(aes(color = porosity))

#combined <- sp1 + sp2 + sp3

#combined

dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = freq))
#dp2 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = factor(chemIndex)))
#dp3 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) + geom_point(aes(color = porosity))

#combined <- dp1 + dp2 + dp3

#combined

combined <- sp1 + dp1
combined
```

In @fig-3, we logged the responses, and plotted them against the log of cycle. The relationships between log(cycle) and the logged responses look more linear than the relationships in @fig-2, so a linear model using these logged variables would fit better.

In order to understand how log(stiffnessFrac) and log(dampingFrac) behave given different frequencies, we also colored the samples in @fig-3 by frequency. For log(stiffnessFrac), we see that higher frequencies have a more negative trend than lower frequencies. For log(dampingFrac), we see that lower frequencies have a more negative trend than higher frequencies. Frequency looks related to both stiffnessFrac and dampingFrac.

```{r}
#| echo: false
#| label: fig-4
#| warning: false
#| fig-cap: "Log(stiffnessFrac) vs Log(cycle), Separated by ChemIndex and Porosity"
#| fig-align: "center"
#| fig-width: 18
#| fig-height: 8
#| 
sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

combined <- sfc1 + sfc2 + sfc3 + sfc4 + sfc5 + sfc6 + plot_layout(ncol = 3)
combined
```

The plots in @fig-4 separate the log(stiffnessFrac) vs log(cycle) graph in @fig-3 to show how the interaction between chemical index and porosity impact log(stiffnessFrac). We can see that the chemical index 79 has many samples where log(stiffnessFrac) trends upwards, whereas chemical indices 100 and 121 mostly trend downwards. This indicates a relationship between log(stiffnessFrac) and chemical index. Downward trends mean the stiffness decreases as cycle increases, which is not what we are looking for in a durable foam.

We also see an effect from porosity. For chemical index 79, low porosity foam trends downward more often than high porosity foam, and for chemical index 100, low porosity foam trends upward more often than high porosity foam. So not only does porosity affect log(stiffnessFrac), but it effects it different ways for different porosities. This suggests that a porosity:chemIndex interaction term would improve our model.

```{r}
#| echo: false
#| label: fig-5
#| warning: false
#| fig-cap: "Log(dampingFrac) vs Log(cycle), Separated by ChemIndex and Porosity"
#| fig-align: "center"
#| fig-width: 18
#| fig-height: 8
#| 
dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

combined <- dfc1 + dfc2 + dfc3 + dfc4 + dfc5 + dfc6 + plot_layout(ncol = 3)
combined
```

The plots in @fig-5 separate the log(dampingFrac) vs log(cycle) graph from above to show how the interaction between chemical index and porosity impact log(dampingFrac). We can see differences in the log(dampingFrac) trends between different chemical indices, so there is likely a relationship present, but this relationship looks less pronounced than it was for the log(stiffnessFrac) trends.

Porosity also appears to have some effect. Trends in log(dampingFrac) for chemical index 121 are more similar to each other for low porosity than for high porosity, and trends in log(dampingFrac) for chemical index 79 are bit more similar to each other for high porosity than for low porosity. This suggests a porosity:chemIndex interaction term would improve our damping model, but this is still less clear than it was for stiffness.

# 3. Statistical Analysis

Based on the initial research questions, we will be developing two separate models: one for stiffness fraction and the other for damping fraction. Due to the experimental design and the fact that our observations are not independent of each other, we used a mixed model format for our predictions.

In order to select our variables, we first started with a mixed model where all variables aside from log(cycle) were treated as level two variables and log(cycle) was treated as a level one variable. In this context, the level two variables on their own will impact the intercept of our model and the interaction between log(cycle) and the other variables will impact the slope of our model. After starting with this full model, we manually removed variables from the model to increase the significance of the coefficients and to minimize the REML. From this process, we were able to create a model using the following predictors:

|    Stiffness Model Predictors    |        Damping Model Predictors                     
|:---------------------------------|:------------------------------------|
           chemIndex               |             chemIndex
             freq                  |               freq
          log(cycle)               |            log(cycle)
              amp                  |             porosity
       chemIndex:log(cycle)        |         chemIndex:log(cycle)
        freq:log(cycle)            |           freq:log(cycle)
        amp:log(cycle)             |         porosity:log(cycle)

\begin{center}
\text{Table 4: Predictors Selected For Each Model}
\end{center}

# 3.1 Stiffness

To best predict the value of log(stiffness fraction), we used the following model:

**Level 1**: $\log(\text{stiffnessFrac}) = \alpha_i + \beta_i*log(cycle)$

**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{chemIndex} + \alpha_2*\text{freq} + \alpha_3*\text{amp}$

$\beta_i = \beta_0 + \beta_1*\text{chemIndex} +\beta_2*\text{freq} + \beta_3*\text{amp}$

From the statistical summary below we see a high significance in our coefficients. The p-value for every term has a significance below five percent, meaning that there is a significant relationship between each of the terms and the response variable. The REML is also extremely low which leads us to conclude that this model fits the data very well and has minimal error. The coefficients not relating to cycle are impacting the initial stiffness of the foam at time zero. The variable that has the most positive impact on the initial stiffness is having a chemical index of 2.

```{r, echo = FALSE}
Anova(FactorModel)
```

We also assessed the model to see how it met the conditions of a predictive model. From @fig-example3 (see Appendix), we see that there appears to be some sinusoidal behavior at the beginning and the middle of the shape has a larger range than the rest of the plot. Since the range of the plot is so small, we can conclude that the equal variance condition is being met. We also do not see any flaring or nonlinear movement so the linearity condition is also being met. The QQ plot in @fig-example4 (see Appendix) shows some tailing on the ends but nothing too concerning. This concludes that all conditions are being met so this model is valid for the purpose of prediction.

Lastly, given that we are using a mixed effects model, it is necessary to explore the random effects associated with the model. From @fig-6, we notice that as the intercept increases, the slope decreases. This is a very defined trend and supports our conclusion that a mixed model should be used. We would expect to see that each of the foams starts at a similar stiffness fraction and as time increases, there would be deviations based on the other predictors. This plot confirms this behavior.

```{r, echo = FALSE}
#| echo: false
#| label: fig-6
#| warning: false
#| fig-cap: "Intercept Rand Effects vs. Slope Rand Effects"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4

re <- ranef(FactorModel)$ID

plot(re)
```

#	3.2 Damping

To best predict the value of log(damping fraction), we used this model:

**Level 1**: $\log(\text{dampingFrac}) = \alpha_i + \beta_i*log(cycle)$

**Level 2**: $\alpha_i = \alpha_0 + \alpha_1*\text{porosity} + \alpha_2*\text{chemIndex} + \alpha_3*\text{freq}$

$\beta_i = \beta_0 + \beta_1*\text{porosity} + \beta_2*\text{chemIndex} +\beta_3*\text{freq}$

The summary of this model shows a high significance in almost all of our coefficients through small p-values (less than 0.05). These small p-values suggest that these variables are significant predictors of our response variable. The only value that does not show a significant p-value is chemIndex, which we chose to keep because the interaction between chemIndex and log(cycle) is significant. This model has multiple interactions: porosity and log(cycle), frequency and log(cycle), and chemical index and log(cycle). These interactions show that if all other variables are held constant, there is a combined effect of these interactions on the response variable.

```{r, echo = FALSE}
Anova(dampingModel)
```

We used a residuals vs fitted plot and QQ plot to assess the conditions of our model (see @fig-example5 and @fig-example6, in the Appendix). We found that this damping model meets all of these conditions which therefore we can analyze our model's results.

Lastly, we must look at the random effects of our model in @fig-7. Like the stiffness model, as the intercept increases, the slope decreases. Although the trend is not as linear as the stiffness model, it still supports our conclusion that a mixed model should be used. Again, we would expect to see that each of the foams starts at a similar damping fraction and as time increases, there would be deviations based on the other predictors. This plot confirms this behavior.

```{r, echo = FALSE}
#| echo: false
#| label: fig-7
#| warning: false
#| fig-cap: "Intercept Rand Effects vs. Slope Rand Effects"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4

re <- ranef(dampingModel)$ID

plot(re)
```

# 4. Recommendations

**Question 1:** We see that chemical index, frequency, and cycle number have the most significant effects on both damping and stiffness. However porosity has an effect on damping and amplitude as an effect on stiffness.

**Question 2:** For stiffness, we see that the individual predictors on their own will increase the initial stiffness, but as time goes on, foam using chemical index 121 will have the greatest decrease in stiffness. The individual predictors will impact damping in a similar way but foam with a chemical index of 79 will see the greatest decay in damping. Overall, there appear to be significant impacts on the stiffness and damping values depending on chemical index and porosity.

# 5. Additional Considerations

One concern we have about our conclusions is that they were made based on relatively few observations. Our model for stiffness fraction, for example, depends on chemical index, frequency, and amplitude, and only 12 samples with chemical index 100 were tested. There are too many possible combinations of frequency and amplitude to be explored thoroughly with just 12 samples. In cases like this, our models' coefficient estimates are based on very little information. The p-values gave us confidence that the coefficients were not zero, but our estimates could still be far off from their true values. We would need more observations to get a more reliable estimate of the true coefficients.

A similar concern comes from foam types and testing conditions that were used less often, such as high frequency. The reliability of our models is already limited, and for these cases they are even less reliable. Additionally, if one variety of foam was never tested at high frequency, then the model's claims about that foam at high frequency are based on nothing, and should not be trusted at all. So not all information this model gives on what effects stiffness and damping in foam is equally trustworthy.

Our models could have also been thrown off by conditions we didn't consider, such as room temperature and how many other people were using the laboratory. Since these conditions were not controlled for, they could have had an effect we were unable to detect.

Another consideration is that our models predict the log of stiffness fraction and the log of damping fraction linearly with the log of cycle number, which is a poor approximation in some cases. It is possible that models that are quadratic in log cycle number could have fit the data with slightly better accuracy, but these models would be harder to interpret. So our final models prioritize understanding the relationship between the responses and the predictors, possibly at the expense of accuracy.

For further studies, we recommend that more foam samples are tested, so that there are more observations to work with. The same number of each type of foam should be used, and the treatments such as amplitude and initial stress should be assigned evenly and randomly, so each treatment level is fully tested on each type of foam. This will allow for more trustworthy models.

# Technical Appendix

### Data Wrangling

```{r}
#| echo: true
#| eval: false
knitr::opts_chunk$set(echo = TRUE)

library(patchwork)
library("readxl")
library(ggplot2)
library(tidyverse)

foamDataRaw <- read_excel("DMA_Fatigue_Data_from_Team_Wendy_Polyurethane
                          _LLNL-MI-860456.xlsx")
names(foamDataRaw) <- foamDataRaw[3,]
foamDataRaw <- foamDataRaw[-c(1:3),] # remove rows without data
foamDataRaw <- foamDataRaw[,-c(62:64)] # remove duplicate columns

foamDataIntermediate <- foamDataRaw |>
  separate(col=SampleCode, into=c("Chemical", "chemIndex", "porosity", "ID")) |>
  select(!Chemical) |>
  mutate_at(vars(chemIndex), as.character) |>
  mutate_at(vars(porosity), as.character) |>
  mutate(chemIndex = dplyr::recode(chemIndex, "79"="1", "100"="2", "121"="3")) |>
  mutate(porosity = dplyr::recode(porosity, "71"="low", "81"="high")) |>
  rename("initialStrain"="strain") |>
  rename("initialStress"="stresskpa") |>
  mutate_at(c("Cyc_A", "Cyc_B", "Cyc_C", "Cyc_D", "Cyc_E", "Cyc_F", "Cyc_G", "Cyc_H", "Cyc_I", 
              "Cyc_J", "Cyc_K", "Cyc_L", "Cyc_M", "Cyc_N", "E1_MPA_Cyc_A", "E1_MPA_Cyc_B", 
              "E1_MPA_Cyc_C", "E1_MPA_Cyc_D", "E1_MPA_Cyc_E", "E1_MPA_Cyc_F", "E1_MPA_Cyc_G", 
              "E1_MPA_Cyc_H", "E1_MPA_Cyc_I", "E1_MPA_Cyc_J", "E1_MPA_Cyc_K", "E1_MPA_Cyc_L", 
              "E1_MPA_Cyc_M", "E1_MPA_Cyc_N", "delta_rad_Cyc_A", "delta_rad_Cyc_B", 
              "delta_rad_Cyc_C", "delta_rad_Cyc_D", "delta_rad_Cyc_E", "delta_rad_Cyc_F", 
              "delta_rad_Cyc_G", "delta_rad_Cyc_H", "delta_rad_Cyc_I", "delta_rad_Cyc_J", 
              "delta_rad_Cyc_K", "delta_rad_Cyc_L", "delta_rad_Cyc_M", "delta_rad_Cyc_N"), 
            as.numeric) |>
  mutate_at(c("E1_MPA_Cyc_A", "E1_MPA_Cyc_B", "E1_MPA_Cyc_C", "E1_MPA_Cyc_D", "E1_MPA_Cyc_E", 
              "E1_MPA_Cyc_F", "E1_MPA_Cyc_G", "E1_MPA_Cyc_H", "E1_MPA_Cyc_I", "E1_MPA_Cyc_J", 
              "E1_MPA_Cyc_K", "E1_MPA_Cyc_L", "E1_MPA_Cyc_M", "E1_MPA_Cyc_N"), abs)

#These lines add columns for Rate Of Change of stiffness and damping.
foamDataIntermediate <- foamDataIntermediate |>
  mutate(stiffness_ROC_A = NA) |>
  mutate(stiffness_ROC_B = (E1_MPA_Cyc_B - E1_MPA_Cyc_A)/(Cyc_B - Cyc_A))|>
  mutate(stiffness_ROC_C = (E1_MPA_Cyc_C - E1_MPA_Cyc_B)/(Cyc_C - Cyc_B))|>
  mutate(stiffness_ROC_D = (E1_MPA_Cyc_D - E1_MPA_Cyc_C)/(Cyc_D - Cyc_C))|>
  mutate(stiffness_ROC_E = (E1_MPA_Cyc_E - E1_MPA_Cyc_D)/(Cyc_E - Cyc_D))|>
  mutate(stiffness_ROC_F = (E1_MPA_Cyc_F - E1_MPA_Cyc_E)/(Cyc_F - Cyc_E))|>
  mutate(stiffness_ROC_G = (E1_MPA_Cyc_G - E1_MPA_Cyc_F)/(Cyc_G - Cyc_F))|>
  mutate(stiffness_ROC_H = (E1_MPA_Cyc_H - E1_MPA_Cyc_G)/(Cyc_H - Cyc_G))|>
  mutate(stiffness_ROC_I = (E1_MPA_Cyc_I - E1_MPA_Cyc_H)/(Cyc_I - Cyc_H))|>
  mutate(stiffness_ROC_J = (E1_MPA_Cyc_J - E1_MPA_Cyc_I)/(Cyc_J - Cyc_I))|>
  mutate(stiffness_ROC_K = (E1_MPA_Cyc_K - E1_MPA_Cyc_J)/(Cyc_K - Cyc_J))|>
  mutate(stiffness_ROC_L = (E1_MPA_Cyc_L - E1_MPA_Cyc_K)/(Cyc_L - Cyc_K))|>
  mutate(stiffness_ROC_M = (E1_MPA_Cyc_M - E1_MPA_Cyc_L)/(Cyc_M - Cyc_L))|>
  mutate(stiffness_ROC_N = (E1_MPA_Cyc_N - E1_MPA_Cyc_M)/(Cyc_N - Cyc_M))|>
  mutate(stiffness_frac_A = E1_MPA_Cyc_A / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_B = E1_MPA_Cyc_B / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_C = E1_MPA_Cyc_C / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_D = E1_MPA_Cyc_D / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_E = E1_MPA_Cyc_E / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_F = E1_MPA_Cyc_F / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_G = E1_MPA_Cyc_G / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_H = E1_MPA_Cyc_H / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_I = E1_MPA_Cyc_I / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_J = E1_MPA_Cyc_J / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_K = E1_MPA_Cyc_K / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_L = E1_MPA_Cyc_L / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_M = E1_MPA_Cyc_M / E1_MPA_Cyc_A) |>
  mutate(stiffness_frac_N = E1_MPA_Cyc_N / E1_MPA_Cyc_A) |>
  mutate(damping_frac_A = tan(delta_rad_Cyc_A) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_B = tan(delta_rad_Cyc_B) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_C = tan(delta_rad_Cyc_C) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_D = tan(delta_rad_Cyc_D) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_E = tan(delta_rad_Cyc_E) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_F = tan(delta_rad_Cyc_F) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_G = tan(delta_rad_Cyc_G) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_H = tan(delta_rad_Cyc_H) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_I = tan(delta_rad_Cyc_I) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_J = tan(delta_rad_Cyc_J) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_K = tan(delta_rad_Cyc_K) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_L = tan(delta_rad_Cyc_L) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_M = tan(delta_rad_Cyc_M) / tan(delta_rad_Cyc_A)) |>
  mutate(damping_frac_N = delta_rad_Cyc_N / delta_rad_Cyc_A) |>mutate(damping_ROC_A = NA) |>
  mutate(damping_ROC_B = (delta_rad_Cyc_B - delta_rad_Cyc_A)/(Cyc_B - Cyc_A))|>
  mutate(damping_ROC_C = (delta_rad_Cyc_C - delta_rad_Cyc_B)/(Cyc_C - Cyc_B))|>
  mutate(damping_ROC_D = (delta_rad_Cyc_D - delta_rad_Cyc_C)/(Cyc_D - Cyc_C))|>
  mutate(damping_ROC_E = (delta_rad_Cyc_E - delta_rad_Cyc_D)/(Cyc_E - Cyc_D))|>
  mutate(damping_ROC_F = (delta_rad_Cyc_F - delta_rad_Cyc_E)/(Cyc_F - Cyc_E))|>
  mutate(damping_ROC_G = (delta_rad_Cyc_G - delta_rad_Cyc_F)/(Cyc_G - Cyc_F))|>
  mutate(damping_ROC_H = (delta_rad_Cyc_H - delta_rad_Cyc_G)/(Cyc_H - Cyc_G))|>
  mutate(damping_ROC_I = (delta_rad_Cyc_I - delta_rad_Cyc_H)/(Cyc_I - Cyc_H))|>
  mutate(damping_ROC_J = (delta_rad_Cyc_J - delta_rad_Cyc_I)/(Cyc_J - Cyc_I))|>
  mutate(damping_ROC_K = (delta_rad_Cyc_K - delta_rad_Cyc_J)/(Cyc_K - Cyc_J))|>
  mutate(damping_ROC_L = (delta_rad_Cyc_L - delta_rad_Cyc_K)/(Cyc_L - Cyc_K))|>
  mutate(damping_ROC_M = (delta_rad_Cyc_M - delta_rad_Cyc_L)/(Cyc_M - Cyc_L))|>
  mutate(damping_ROC_N = (delta_rad_Cyc_N - delta_rad_Cyc_M)/(Cyc_N - Cyc_M))

# These lines lengthen the data by putting each measurement interval in its own data frame, 
#and row-binding them together.
foamDataA <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_A, E1_MPA_Cyc_A, 
                    stiffness_ROC_A, stiffness_frac_A, delta_rad_Cyc_A, damping_ROC_A, 
                    damping_frac_A, NVP_Cyc_A) |>
  rename("cycle"="Cyc_A") |>
  rename("stiffness"="E1_MPA_Cyc_A") |>
  rename("damping"="delta_rad_Cyc_A") |>
  rename("NVP"="NVP_Cyc_A") |>
  mutate("intervalCode"="A") |>
  rename("stiffnessROC"="stiffness_ROC_A") |>
  rename("stiffnessFrac"="stiffness_frac_A") |>
  rename("dampingROC"="damping_ROC_A") |>
  rename("dampingFrac"="damping_frac_A")
foamDataB <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_B, E1_MPA_Cyc_B, 
                    stiffness_ROC_B, stiffness_frac_B, delta_rad_Cyc_B, damping_ROC_B, 
                    damping_frac_B, NVP_Cyc_B) |>
  rename("cycle"="Cyc_B") |>
  rename("stiffness"="E1_MPA_Cyc_B") |>
  rename("damping"="delta_rad_Cyc_B") |>
  rename("NVP"="NVP_Cyc_B") |>
  mutate("intervalCode"="B") |>
  rename("stiffnessROC"="stiffness_ROC_B") |>
  rename("stiffnessFrac"="stiffness_frac_B") |>
  rename("dampingROC"="damping_ROC_B") |>
  rename("dampingFrac"="damping_frac_B")
foamDataC <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_C, E1_MPA_Cyc_C, 
                    stiffness_ROC_C, stiffness_frac_C, delta_rad_Cyc_C, damping_ROC_C, 
                    damping_frac_C, NVP_Cyc_C) |>
  rename("cycle"="Cyc_C") |>
  rename("stiffness"="E1_MPA_Cyc_C") |>
  rename("damping"="delta_rad_Cyc_C") |>
  rename("NVP"="NVP_Cyc_C") |>
  mutate("intervalCode"="C") |>
  rename("stiffnessROC"="stiffness_ROC_C") |>
  rename("stiffnessFrac"="stiffness_frac_C") |>
  rename("dampingROC"="damping_ROC_C") |>
  rename("dampingFrac"="damping_frac_C")
foamDataD <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_D, E1_MPA_Cyc_D, 
                    stiffness_ROC_D, stiffness_frac_D, delta_rad_Cyc_D, damping_ROC_D, 
                    damping_frac_D, NVP_Cyc_D) |>
  rename("cycle"="Cyc_D") |>
  rename("stiffness"="E1_MPA_Cyc_D") |>
  rename("damping"="delta_rad_Cyc_D") |>
  rename("NVP"="NVP_Cyc_D") |>
  mutate("intervalCode"="D") |>
  rename("stiffnessROC"="stiffness_ROC_D") |>
  rename("stiffnessFrac"="stiffness_frac_D") |>
  rename("dampingROC"="damping_ROC_D") |>
  rename("dampingFrac"="damping_frac_D")
foamDataE <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_E, E1_MPA_Cyc_E, 
                    stiffness_ROC_E, stiffness_frac_E, delta_rad_Cyc_E, damping_ROC_E, 
                    damping_frac_E, NVP_Cyc_E) |>
  rename("cycle"="Cyc_E") |>
  rename("stiffness"="E1_MPA_Cyc_E") |>
  rename("damping"="delta_rad_Cyc_E") |>
  rename("NVP"="NVP_Cyc_E") |>
  mutate("intervalCode"="E") |>
  rename("stiffnessROC"="stiffness_ROC_E") |>
  rename("stiffnessFrac"="stiffness_frac_E") |>
  rename("dampingROC"="damping_ROC_E") |>
  rename("dampingFrac"="damping_frac_E")
foamDataF <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_F, E1_MPA_Cyc_F, 
                    stiffness_ROC_F, stiffness_frac_F, delta_rad_Cyc_F, damping_ROC_F, 
                    damping_frac_F, NVP_Cyc_F) |>
  rename("cycle"="Cyc_F") |>
  rename("stiffness"="E1_MPA_Cyc_F") |>
  rename("damping"="delta_rad_Cyc_F") |>
  rename("NVP"="NVP_Cyc_F") |>
  mutate("intervalCode"="F") |>
  rename("stiffnessROC"="stiffness_ROC_F") |>
  rename("stiffnessFrac"="stiffness_frac_F") |>
  rename("dampingROC"="damping_ROC_F") |>
  rename("dampingFrac"="damping_frac_F")
foamDataG <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_G, E1_MPA_Cyc_G, 
                    stiffness_ROC_G, stiffness_frac_G, delta_rad_Cyc_G, damping_ROC_G, 
                    damping_frac_G, NVP_Cyc_G) |>
  rename("cycle"="Cyc_G") |>
  rename("stiffness"="E1_MPA_Cyc_G") |>
  rename("damping"="delta_rad_Cyc_G") |>
  rename("NVP"="NVP_Cyc_G") |>
  mutate("intervalCode"="G") |>
  rename("stiffnessROC"="stiffness_ROC_G") |>
  rename("stiffnessFrac"="stiffness_frac_G") |>
  rename("dampingROC"="damping_ROC_G") |>
  rename("dampingFrac"="damping_frac_G")
foamDataH <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_H, E1_MPA_Cyc_H, 
                    stiffness_ROC_H, stiffness_frac_H, delta_rad_Cyc_H, damping_ROC_H, 
                    damping_frac_H, NVP_Cyc_H) |>
  rename("cycle"="Cyc_H") |>
  rename("stiffness"="E1_MPA_Cyc_H") |>
  rename("damping"="delta_rad_Cyc_H") |>
  rename("NVP"="NVP_Cyc_H") |>
  mutate("intervalCode"="H") |>
  rename("stiffnessROC"="stiffness_ROC_H") |>
  rename("stiffnessFrac"="stiffness_frac_H") |>
  rename("dampingROC"="damping_ROC_H") |>
  rename("dampingFrac"="damping_frac_H")
foamDataI <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_I, E1_MPA_Cyc_I, 
                    stiffness_ROC_I, stiffness_frac_I, delta_rad_Cyc_I, damping_ROC_I, 
                    damping_frac_I, NVP_Cyc_I) |>
  rename("cycle"="Cyc_I") |>
  rename("stiffness"="E1_MPA_Cyc_I") |>
  rename("damping"="delta_rad_Cyc_I") |>
  rename("NVP"="NVP_Cyc_I") |>
  mutate("intervalCode"="I") |>
  rename("stiffnessROC"="stiffness_ROC_I") |>
  rename("stiffnessFrac"="stiffness_frac_I") |>
  rename("dampingROC"="damping_ROC_I") |>
  rename("dampingFrac"="damping_frac_I")
foamDataJ <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_J, E1_MPA_Cyc_J, 
                    stiffness_ROC_J, stiffness_frac_J, delta_rad_Cyc_J, damping_ROC_J, 
                    damping_frac_J, NVP_Cyc_J) |>
  rename("cycle"="Cyc_J") |>
  rename("stiffness"="E1_MPA_Cyc_J") |>
  rename("damping"="delta_rad_Cyc_J") |>
  rename("NVP"="NVP_Cyc_J") |>
  mutate("intervalCode"="J") |>
  rename("stiffnessROC"="stiffness_ROC_J") |>
  rename("stiffnessFrac"="stiffness_frac_J") |>
  rename("dampingROC"="damping_ROC_J") |>
  rename("dampingFrac"="damping_frac_J")
foamDataK <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_K, E1_MPA_Cyc_K, 
                    stiffness_ROC_K, stiffness_frac_K, delta_rad_Cyc_K, damping_ROC_K, 
                    damping_frac_K, NVP_Cyc_K) |>
  rename("cycle"="Cyc_K") |>
  rename("stiffness"="E1_MPA_Cyc_K") |>
  rename("damping"="delta_rad_Cyc_K") |>
  rename("NVP"="NVP_Cyc_K") |>
  mutate("intervalCode"="K") |>
  rename("stiffnessROC"="stiffness_ROC_K") |>
  rename("stiffnessFrac"="stiffness_frac_K") |>
  rename("dampingROC"="damping_ROC_K") |>
  rename("dampingFrac"="damping_frac_K")
foamDataL <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_L, E1_MPA_Cyc_L, 
                    stiffness_ROC_L, stiffness_frac_L, delta_rad_Cyc_L, damping_ROC_L, 
                    damping_frac_L, NVP_Cyc_L) |>
  rename("cycle"="Cyc_L") |>
  rename("stiffness"="E1_MPA_Cyc_L") |>
  rename("damping"="delta_rad_Cyc_L") |>
  rename("NVP"="NVP_Cyc_L") |>
  mutate("intervalCode"="L") |>
  rename("stiffnessROC"="stiffness_ROC_L") |>
  rename("stiffnessFrac"="stiffness_frac_L") |>
  rename("dampingROC"="damping_ROC_L") |>
  rename("dampingFrac"="damping_frac_L")
foamDataM <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_M, E1_MPA_Cyc_M, 
                    stiffness_ROC_M, stiffness_frac_M, delta_rad_Cyc_M, damping_ROC_M, 
                    damping_frac_M, NVP_Cyc_M) |>
  rename("cycle"="Cyc_M") |>
  rename("stiffness"="E1_MPA_Cyc_M") |>
  rename("damping"="delta_rad_Cyc_M") |>
  rename("NVP"="NVP_Cyc_M") |>
  mutate("intervalCode"="M") |>
  rename("stiffnessROC"="stiffness_ROC_M") |>
  rename("stiffnessFrac"="stiffness_frac_M") |>
  rename("dampingROC"="damping_ROC_M") |>
  rename("dampingFrac"="damping_frac_M")
foamDataN <- select(foamDataIntermediate, chemIndex:initialStress, Cyc_N, E1_MPA_Cyc_N, 
                    stiffness_ROC_N, stiffness_frac_N, delta_rad_Cyc_N, damping_ROC_N, 
                    damping_frac_N, NVP_Cyc_N) |>
  rename("cycle"="Cyc_N") |>
  rename("stiffness"="E1_MPA_Cyc_N") |>
  rename("damping"="delta_rad_Cyc_N") |>
  rename("NVP"="NVP_Cyc_N") |>
  mutate("intervalCode"="N") |>
  rename("stiffnessROC"="stiffness_ROC_N") |>
  rename("stiffnessFrac"="stiffness_frac_N") |>
  rename("dampingROC"="damping_ROC_N") |>
  rename("dampingFrac"="damping_frac_N")

foamData <- rbind(foamDataA, foamDataB, foamDataC, foamDataD, foamDataE, foamDataF, foamDataG, 
                  foamDataH, foamDataI, foamDataJ, foamDataK, foamDataL, foamDataM, foamDataN)

# Filter rows with NA in the cycle, stiffness, and damping columns
foamData <- filter(foamData, !is.na(cycle))

# Reorder the columns
foamData <- select(foamData, 3, 1, 2, 4, 5, 6, 7, 16, 8, 11, 14, 9, 12, 10, 13, 15)

# write the file
write_csv(foamData, "foamData.csv")
```

### R Script for plots, models, and tables in Sections 2 and 3:

```{r}
#| echo: true
#| eval: false

foam_data <- read.csv("foamData.csv")

#Some extra wrangling
foam_data_scaled <- foam_data |>
  filter(cycle <= 10000 & cycle > 0)

FactorModel <- lmer(log(stiffnessFrac) ~  porosity + factor(chemIndex) 
                    + factor(chemIndex):porosity + log(cycle) +  factor(chemIndex):log(cycle) 
                    + porosity:log(cycle) + factor(chemIndex):porosity:log(cycle) 
                    + (1 + log(cycle)|ID), data = foam_data_scaled)

dampingModel <- lmer(log(dampingFrac) ~ porosity + factor(chemIndex) + freq + log(cycle) 
                     + porosity:log(cycle) + freq:log(cycle) + factor(chemIndex):log(cycle) 
                     + (1 + log(cycle)|ID), data = foam_data_scaled)

# Convert intervalCode to a factor with specified levels
foam_data_scaled$intervalCode <- factor(foam_data_scaled$intervalCode, levels = c("A", "B", "C", 
                                       "D","E", "F", "G", "H","I", "J", "K", "L","M", "N"))

# EDA code

# graph1
foam_Samples <- foam_data[ ,c("ID", "chemIndex", "porosity", "freq", "amp", "initialStrain", 
                              "initialStress")]
foam_Samples <- unique(foam_Samples)

count(foam_Samples, chemIndex, porosity)

ggplot(foam_Samples, aes(x=freq)) +
  geom_histogram(bins=10)

# graph2
p1 <- ggplot(foam_data_scaled, aes(y = stiffnessFrac, x = cycle, col=factor(ID))) +
  geom_point() + 
  theme(legend.position="none")

p2 <- ggplot(foam_data_scaled, aes(y = dampingFrac, x = cycle, col=factor(ID))) + 
  geom_point() +
  theme(legend.position="none")

combined <- p1 + p2

combined

# graph3
sp1 <- ggplot(foam_data_scaled, aes(y = log(stiffnessFrac), x = log(cycle))) 
+ geom_point(aes(color = freq))

dp1 <- ggplot(foam_data_scaled, aes(y = log(dampingFrac), x = log(cycle))) 
+ geom_point(aes(color = freq))

combined <- sp1 + dp1
combined

# graph4
sfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity 
                                == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity 
                                == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity 
                                == "high",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
sfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data$porosity 
                                == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data$porosity
                                == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
sfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data$porosity
                                == "low",], aes(x=log(cycle), y=log(stiffnessFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.325, 0.15) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

combined <- sfc1 + sfc2 + sfc3 + sfc4 + sfc5 + sfc6 + plot_layout(ncol = 3)
combined

# graph5
dfc1 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity 
                                == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc2 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity 
                                == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc3 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity 
                                == "high",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="cyan2", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 81") +
  theme(plot.title = element_text(size=25))
dfc4 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 1 & foam_data_scaled$porosity 
                                == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="red3", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 79, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc5 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 2 & foam_data_scaled$porosity 
                                == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(color="green4", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 100, Porisity 71") +
  theme(plot.title = element_text(size=25))
dfc6 <- ggplot(foam_data_scaled[foam_data_scaled$chemIndex == 3 & foam_data_scaled$porosity 
                                == "low",], aes(x=log(cycle), y=log(dampingFrac))) +
  geom_point(col="blue", size=3) + 
  geom_hline(yintercept = 0, linetype="dashed") +
  ylim(-0.175, 0.125) +
  labs(title="ChemIndex 121, Porisity 71") +
  theme(plot.title = element_text(size=25))

combined <- dfc1 + dfc2 + dfc3 + dfc4 + dfc5 + dfc6 + plot_layout(ncol = 3)
combined

Anova(FactorModel)

# random effects of stiffnessFrac model
re <- ranef(FactorModel)$ID

plot(re)

Anova(dampingModel)

# random effects of dampingFrac model
re <- ranef(dampingModel)$ID

plot(re)
```

```{r}
#| echo: true
#| label: fig-example1
#| warning: false
#| fig-cap: "Backward Selection for stiffnessFrac Predictors"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4
regfit.full <- regsubsets(log(stiffnessFrac) ~ chemIndex + porosity + freq + amp 
                          + initialStrain + initialStress + log(cycle) + (chemIndex 
                          + porosity + freq + amp + initialStrain + initialStress)*log(cycle) 
                          + (1 + log(cycle)|ID), 
                          data = foam_data_scaled)
#regfit.full <- regsubsets(stiffnessFrac ~ ., data = foam_data_scaled)
plot(regfit.full, scale = 'bic')
```

```{r}
#| echo: true
#| label: fig-example2
#| warning: false
#| fig-cap: "Backward Selection for dampingFrac Predictors"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4
#| 
regfit.full <- regsubsets(log(dampingFrac) ~ chemIndex + porosity + freq + amp 
                          + initialStrain + initialStress + log(cycle) + chemIndex:porosity 
                          + (chemIndex + porosity + freq + amp + initialStrain 
                          + initialStress + chemIndex:porosity)*log(cycle) 
                          + (1 + log(cycle)|ID), data = foam_data_scaled)

plot(regfit.full, scale = "bic")
```

```{r}
#| echo: true
#| label: fig-example3
#| warning: false
#| fig-cap: "Residual Plot of Stiffness Fraction Model"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4

plot(FactorModel)
```

```{r}
#| echo: true
#| label: fig-example4
#| warning: false
#| fig-cap: "QQ Plot of Stiffness Fraction Model"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4

qqnorm(resid(FactorModel))
qqline(resid(FactorModel))
```

```{r}
#| echo: true
#| label: fig-example5
#| warning: false
#| fig-cap: "Residual Plot of Damping Fraction Model"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4

plot(dampingModel)
```

```{r}
#| echo: true
#| label: fig-example6
#| warning: false
#| fig-cap: "QQ Plot of Damping Fraction Model"
#| fig-align: "center"
#| fig-width: 7
#| fig-height: 4
qqnorm(resid(dampingModel))
qqline(resid(dampingModel))
```